### 2022-11-18 response.tex

\begin{center}
	\large
	Response Letter
\end{center}
We thank the reviewers for their useful comments and feedback on improving the manuscript and the presentation. In the following, we summarize changes reflected in the revised manuscript.

\paragraph{Distribution shifts in life-long learning.} We propose Hamming distance heuristics in {\imli} with the assumption that distribution remains same. One way to account for distribution shift is to consider last $ p $ batches instead of the (single) previous batch in the objective function in mini-batch learning. For a feature variable $ B^i_j $, we consider its majority assignment in last $ p $ classification rules and encode as a soft clause to retain the majority assignment in the current batch. Moreover, we can reweigh the soft clause by prioritizing assignments of $ B^i_j $ in recent batches.
%When distribution changes, the classifier $ \mathcal{R}_{i} $ for the $ i^ $ 
\red{[Add page number]}

\paragraph{Discussions of related works in experiments.}{\imli} scales well than existing interpretable rule-based classifiers because of the incremental solving approach. Most existing works rely on rule-mining of potential classification rules followed by an optimization algorithm such as Bayesian optimization and branch and bound algorithm. We observe incrementality of learning to be more effective. We elaborate this in Section~\ref{interpretability_imli_sec:experiments_scalability}, page 39.

\paragraph{How does {\justicia} outperform a direct approach by learning a Bayesian Network on limited samples?} We agree with the reviewer that learning a Bayesian Network on limited samples does not always result in a better robustness (e.g., less standard deviation of fairness metrics estimation) of {\justicia} than the direct approach estimating fairness on a dataset.

Elaborately, in Chapter~\ref{chapter:justicia} Figure~\ref{fairness_justicia_fig:sample-size}, {\justicia} demonstrates higher robustness than the direct approach, where we consider the distribution in two ways: (i) non-sensitive features are independent and (ii) they are dependent on sensitive features only. Thus, Figure~\ref{fairness_justicia_fig:sample-size} does not involve experiments with Bayesian Network, which we introduce later in Chapter~\ref{chapter:fvgm}. In our revised experiment, we observe that {\justicia} with learning Bayesian network on a dataset, called {\fvgm}, exhibits less robustness due to network learning.

\paragraph{Clarification on Fairness Influence Function (FIF).} We consider an axiom that FIFs are additive where the sum of FIFs of all subsets of non-sensitive features is equal \textit{to the resultant unfairness of the classifier}. Here the unfairness of the classifier is a real value in $ [0,1] $, such as statistical parity, equalized odds, and predictive parity. While one may choose a different axiom, such as multiplicative FIFs, our choice is motivated by the explainability literature for machine learning, such as in SHAP~\cite{begley2020explainability,lundberg2020explaining}. In particular, additivity of FIFs is based on the idea of decomposing the bias/unfairness as the sum of the contribution of different subsets of features on the unfairness of the classifier.

### 2022-12-02 response.tex

\begin{center}
	\large
	Response Letter
\end{center}
We thank the reviewers for their useful comments and feedback on improving the manuscript and the presentation. In the following, we summarize changes reflected in the revised manuscript.

\paragraph{Distribution shifts in life-long learning.} We propose Hamming distance heuristics in {\imli} with the assumption that distribution remains same across batches. One way to account for distribution shifts is to consider last $ p $  $ (> 1) $ batches instead of the (single) previous batch in the objective function in mini-batch learning. For a feature variable $ B^i_j $, we consider its majority assignment in last $ p $ classification rules and encode as a soft clause to retain the majority assignment in the current batch. Moreover, we can reweigh the soft clause by prioritizing assignments of $ B^i_j $ in recent batches. We mention this in Section~\ref{interpretability_imli_sec:incremental_learning}, page 28.
%When distribution changes, the classifier $ \mathcal{R}_{i} $ for the $ i^ $

\paragraph{Discussions of related works in experiments.}{\imli} scales well than existing interpretable rule-based classifiers because of the incremental solving approach. Most existing works rely on rule-mining of potential classification rules followed by an optimization algorithm such as Bayesian optimization and branch and bound algorithm. In experiments, we observe the incremental learning of {\imli} to be more scalable  than the state-of-the-art methods. We elaborate this in Section~\ref{interpretability_imli_sec:experiments_scalability}, page 39.

\paragraph{How does {\justicia} outperform a direct approach by learning a Bayesian Network on limited samples?} We agree with the reviewer that learning a Bayesian Network on limited samples does not always result in a better robustness (e.g., less standard deviation of fairness metrics estimation) of {\justicia} than the direct approach of estimating fairness on a dataset.

Elaborately, in Chapter~\ref{chapter:justicia} Figure~\ref{fairness_justicia_fig:sample-size}, {\justicia} demonstrates higher robustness than the direct approach, where we consider a specific distribution of non-sensitive features conditioning only on sensitive features. Thus, Figure~\ref{fairness_justicia_fig:sample-size} does not involve experiments with Bayesian Network capturing correlations of all features, which we indeed introduce later in Chapter~\ref{chapter:fvgm}. In our revised experiment, we observe that {\justicia} with Bayesian network, called {\fvgm}, exhibits less robustness due to Bayesian network learning.

\paragraph{Clarification on Fairness Influence Function (FIF).} Our additive axiom for FIFs is based on the idea of decomposing the total unfairness of the classifiers among different subsets of features~\cite{begley2020explainability,lundberg2020explaining}. We consider that the sum of FIFs of all subsets of non-sensitive features is equal \textit{to the resultant unfairness of the classifier}, where unfairness is a real number in $ [0,1] $, such as statistical parity, equalized odds, and predictive parity.

### 2022-12-02 chapters/background.tex
adding appendix

\subsection{Bayesian Network}
In general, a Probabilistic Graphical Model~\cite{koller2009probabilistic}, and specifically a \textit{Bayesian network}~\cite{pearl1985bayesian,chavira2008probabilistic}, encodes the dependencies and conditional independence between a set of random variables. In fairness verification with correlated features, we leverage an access to a Bayesian network on sensitive and non-sensitive features $ \nonsensitive \cup \sensitive $ that represents the joint distribution on them. A Bayesian network is denoted by a pair $ (\graph, \theta)$, where $ \graph \triangleq (\graphvertices, \graphedges) $ is a DAG (Directed Acyclic Graph), and $\theta$ is a set of parameters encoding the conditional probabilities induced by the joint distribution under investigation. Each vertex $\graphvertex_i \in \graphvertices$ corresponds to a random variable. Edges $ \graphedges \in \graphvertices \times \graphvertices $ imply conditional dependencies among variables. For each variable $ \graphvertex_i \in \graphvertices $, let $ \parent(\graphvertex_i) \subseteq \graphvertices \setminus \{\graphvertex_i\} $ denote the set of parents of $ \graphvertex_i $. Given $\parent(\graphvertex_i)$ and parameters $\theta$, $ \graphvertex_i $ is independent of its other non-descendant variables in $\graph$. Thus, for the assignment $ v_i $ of $ \graphvertex_i $ and $ \mathbf{u} $ of $ \parent(\graphvertex_i) $, the aforementioned semantics of a Bayesian network encodes the joint distribution of $\graphvertices$ as:

>>>>

\subsection{Bayesian Network}
\label{chapter_fairness_preliminaries_BN}
In general, a Probabilistic Graphical Model~\cite{koller2009probabilistic}, and specifically a \textit{Bayesian network}~\cite{pearl1985bayesian,chavira2008probabilistic}, encodes the dependencies and conditional independence between a set of random variables. In fairness verification with correlated features, we leverage an access to a Bayesian network on sensitive and non-sensitive features $ \nonsensitive \cup \sensitive $ that represents the joint distribution on them. A Bayesian network is denoted by a pair $ (\graph, \theta)$, where $ \graph \triangleq (\graphvertices, \graphedges) $ is a DAG (Directed Acyclic Graph), and $\theta$ is a set of parameters encoding the conditional probabilities induced by the joint distribution under investigation. Each vertex $\graphvertex_i \in \graphvertices$ corresponds to a random variable. Edges $ \graphedges \in \graphvertices \times \graphvertices $ imply conditional dependencies among variables. For each variable $ \graphvertex_i \in \graphvertices $, let $ \parent(\graphvertex_i) \subseteq \graphvertices \setminus \{\graphvertex_i\} $ denote the set of parents of $ \graphvertex_i $. Given $\parent(\graphvertex_i)$ and parameters $\theta$, $ \graphvertex_i $ is independent of its other non-descendant variables in $\graph$. Thus, for the assignment $ v_i $ of $ \graphvertex_i $ and $ \mathbf{u} $ of $ \parent(\graphvertex_i) $, the aforementioned semantics of a Bayesian network encodes the joint distribution of $\graphvertices$ as:

### 2022-12-02 chapters/interpretability/imli/incremental_learning.tex
adding appendix

\paragraph{A Heuristic for Mini-Batch Learning.} Let $ \Rule' $ be a classifier learned on the previous batch. In mini-batch learning, we aim to learn a new classifier $ \Rule $ that can generalize to both the current batch and previously seen samples. For that, we consider a soft constraint such that $ \Rule $ does not \emph{differ much} from $ \Rule' $ while training on the current batch. Thus, we hypothesize that by constraining $ \Rule $ to be syntactically similar to $ \Rule' $, it is possible to generalize well; because samples in all batches originate from the same distribution. Since our study focuses on rule-based classifiers, we consider the Hamming distance between two classifiers as a notion of their syntactic dissimilarity. In the following, we define the Hamming distance between two CNF classifiers $ \Rule $, $ \Rule' $ with the same number of clauses as follows.
\[
	d_{H}(\Rule, \Rule') = \sum_{i=1}^k \Big(\sum_{v\in C_i} \mathds{1}(v \not\in C'_i) +  \sum_{v\in C'_i} \mathds{1}(v \not\in C_i) \Big), 
\]

>>>>

\paragraph{A Heuristic for Mini-Batch Learning.} Let $ \Rule' $ be a classifier learned on the previous batch. In mini-batch learning, we aim to learn a new classifier $ \Rule $ that can generalize to both the current batch and previously seen samples. For that, we consider a soft constraint such that $ \Rule $ does not \emph{differ much} from $ \Rule' $ while training on the current batch. Thus, we hypothesize that by constraining $ \Rule $ to be syntactically similar to $ \Rule' $, it is possible to generalize well; because samples in all batches originate from the same distribution\footnote{\blue{We propose Hamming distance heuristics in {\imli} with the assumption that distribution remains same across batches. One way to account for distribution shifts is to consider last $ p $  $ (> 1) $ batches instead of the (single) previous batch in the objective function in mini-batch learning. For a feature variable $ B^i_j $, we consider its majority assignment in last $ p $ classification rules and encode as a soft clause to retain the majority assignment in the current batch. Moreover, we can reweigh the soft clause by prioritizing assignments of $ B^i_j $ in recent batches.}}. Since our study focuses on rule-based classifiers, we consider the Hamming distance between two classifiers as a notion of their syntactic dissimilarity. In the following, we define the Hamming distance between two CNF classifiers $ \Rule $, $ \Rule' $ with the same number of clauses as follows.
\[
	d_{H}(\Rule, \Rule') = \sum_{i=1}^k \Big(\sum_{v\in C_i} \mathds{1}(v \not\in C'_i) +  \sum_{v\in C'_i} \mathds{1}(v \not\in C_i) \Big), 
\]
